Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-02-23 pkg/disttree/inst/draft/draft.Rnw
small corrections and additional reference for Breiman and Cutler's random forests
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2038 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 81c7c29c962330b2090d54d5fec5e755ea6ef090

In particular, 14 basic forecast variables are considered with up 
to 12 variations of each of these variables being included as 
covariates. Since an ensemble of individual predictions is provided 
for each forecast variable, the possible variations include the mean, 
the standard deviation, the minimum and the maximum over all ensemble 
members, noted as ensemble mean, ensemble std. deviation, ensemble 
minimum and ensemble maximum. All of these values can be calculated 
on the sums over 24 hours (6--6 UTC of the following day) or a 6 hour 
time window. Another possible option is to consider the mean, minimum 
or maximum over 24 hours for each individual prediction and then take 
the ensemble mean or ensemble standard deviation, noted as ensemble 
mean of mean/minimal/maximal values or ensemble std. deviation of 
mean/minimal/maximal values respectively. Table~\ref{covariates} 
shows all variables together with the number of variations and 
the provided variations listed in the last column. Altogether 
82~covariates are available in this setting. 
\fixme{Mention that dswrf\_mean/sprd\_min is not included?}

@@

\item An ensemble model output statistics 
(EMOS, \citealp{Gneiting+Raftery+Westveld:2005}) model where the 
location parameter and the logarithm of the scale parameter depend 
linearly on the ensemble mean forecast of the total daily precipitation 
sum and the corresponding ensemble standard deviation respectively.
The choice of the logarithmic function as link function for the scale 
parameter has been made based on the results of 
\cite{Gebetsberger+Messner+Mayr:2017} where different link functions 
have been compared. Moreover, testing the identity, quadratic and 
logarithmic function as link functions in this setting supported 
this choice as the logarithmic function led to the best performance 
of the EMOS model in this application.
\item A GAMLSS where for each distribution parameter the selection 
of covariates and the specification of interactions is done in 
advance, both based on meteorological expert knowledge following 
\cite{Stauffer+Umlauf+Messner:2017} (prespecified GAM). Among the 
available 82 regressors, 9 where selected for the location parameter 
and 6 for the scale parameter. Additionally, for each distribution 
parameter one interaction term is specified in which convective 
available energy is interacting with total precipitation (with the 
ensemble mean for the location parameter and with the ensemble 
standard deviation for the scale parameter). These interactions 
might capture events with a high potential for moderate or severe 
thunderstorms with large amounts of precipitation. By putting 
extra effort in the specification of the formulas the prespecified 
GAM is provided with additional information and is therefore 
expected to profit from this advantage.
\item A boosted GAMLSS \citep{Hofner+Mayr+Schmid:2016} using 
non-homogeneous boosting \citep{Messner+Mayr+Zeileis:2017} 
for an automatic selection of covariates (boosted GAM). 
The idea of this boosting approach is to iteratively minimize 
a loss function where in each step only the coefficient leading 
to the highest improvement of the current model is updated. As 
the initial coefficients are all zero, only the most important 
covariates are included with nonzero coefficients while less 
relevant variables are ignored if the procedure is stopped before 
convergence.
\end{itemize}

@@

The CRPS measures the accuracy of a probabilistic model. 
For one observation $y$ with the predicted distribution 
function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function.
%%% <TH> y > z??? </TH>

In this application the CRPS is evaluated out of sample, 
e.g. on a part of the data set that has not been used to 
build the model. The mean CRPS over this subset of the data 
is used for validation. Since the CRPS can be seen as a 
generalization of the mean absolute error, a low value indicates 
a good fit. In applications, one big advantage in terms of 
interpretability is that the resulting value is on the scale of 
the observations.

@@

After analyzing the results of the models on only one data set where 
visualization is easier, we now want to get a more profound impression 
of the performances of the different methods. For that reason the four 
methods are evaluated in a cross validation framework for the station Axams. 
The data set over all 28 years is split into 7 subsets each consisting of 
4 randomly selected years. While 6 of these subsets are used as learning 
data the 7th is the testing data set for which predictions are made and 
evaluated. This is done by calculating the CRPS value for each of the 
observations in the testing data set and then averaging them. Each of the 
7 parts is selected once as testing data which leads to 7 different settings 
of which again the average values are stored. This whole procedure is repeated 
10 times. The resulting skill scores are illustrated in 
Figure~\ref{boxplots_crps_axams} using the EMOS model as the reference model 
represented by the horizontal line at height 0. This CRPS skill score boxplot 
shows that, in this setting, both GAMs and the distributional forest perform 
better than the EMOS model. While the two GAMs lead to an improvement of 
around 4 percent with the boosted version being slightly ahead, the 
distributional forest even reaches an improvement rate close to 6 percent.
<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams_7x10.rda")){
  load("rain_Axams_7x10.rda")
} else {
  source("rain_cross_7x10.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)

>>>>

In particular, 14 basic forecast variables are considered with up 
to 12 variations of each of these variables being included as 
covariates. Since an ensemble of individual predictions is provided 
for each forecast variable, the possible variations include the mean, 
the standard deviation, the minimum and the maximum over all ensemble 
members, noted as ensemble mean, ensemble std. deviation, ensemble 
minimum and ensemble maximum. All of these values can be calculated 
on the sums over 24 hours (6--6 UTC of the following day) or a 6 hour 
time window. Another possible option is to consider the mean, minimum 
or maximum over 24 hours for each individual prediction and then take 
the ensemble mean or ensemble standard deviation, noted as ensemble 
mean of mean/minimal/maximal values or ensemble std. deviation of 
mean/minimal/maximal values respectively. Table~\ref{covariates} 
shows all variables together with the number of variations and 
the provided variations listed in the last column. Altogether 
82~covariates are available in this setting.

\fixme{Mention that dswrf\_mean/sprd\_min is not included?}

@@

\item An ensemble model output statistics 
(EMOS, \citealp{Gneiting+Raftery+Westveld:2005}) model where the 
location parameter and the logarithm of the scale parameter depend 
linearly on the ensemble mean forecast of total daily precipitation 
sums and the corresponding ensemble standard deviation respectively.
The choice of the logarithmic function as link function for the scale 
parameter has been made based on the results of 
\cite{Gebetsberger+Messner+Mayr:2017} where different link functions 
have been compared. Moreover, testing the identity, quadratic and 
logarithmic function as link functions in this setting supported 
this choice as the logarithmic function led to the best performance 
of the EMOS model in this application.
\item A GAMLSS where for each distribution parameter the selection 
of covariates and the specification of interactions is done in 
advance, both based on meteorological expert knowledge following 
\cite{Stauffer+Umlauf+Messner:2017} (prespecified GAM). Among the 
available 82 regressors, 9 where selected for the location parameter 
and 6 for the scale parameter. Additionally, for each distribution 
parameter one interaction term is specified in which convective 
available energy is interacting with total precipitation (with the 
ensemble mean for the location parameter and with the ensemble 
standard deviation for the scale parameter). These interactions 
might capture events with a high potential for moderate or severe 
thunderstorms with large amounts of precipitation. By putting 
extra effort in the specification of the formulas the prespecified 
GAM is provided with additional information and is therefore 
expected to profit from this advantage.
\item A boosted GAMLSS \citep{Hofner+Mayr+Schmid:2016} using 
non-homogeneous boosting \citep{Messner+Mayr+Zeileis:2017} 
for an automatic selection of covariates (boosted GAM). 
The idea of this boosting approach is to iteratively minimize 
a loss function where in each step only the coefficient leading 
to the highest improvement of the current model is updated. As 
the initial coefficients are all zero, only the most important 
covariates are included with nonzero coefficients while less 
relevant variables are ignored if the procedure is stopped before 
convergence.
\end{itemize}

@@

The CRPS measures the accuracy of a probabilistic model. 
For one observation $y$ with predicted distribution 
function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function.
%%% <TH> y > z??? </TH>

In this application the CRPS is evaluated out of sample, 
e.g. on a part of the data set that has not been considered to 
build the model. The mean CRPS over this subset of the data 
is used for validation. Since the CRPS can be seen as a 
generalization of the mean absolute error, a low value indicates 
a good fit. In applications, one big advantage in terms of 
interpretability is that the resulting value is on the scale of 
the observations.

@@

After analyzing the results of the models on only one data set where 
visualization is easier, the methods are now going to be evaluated
in a cross validation framework for the station Axams. 
The data set over all 28 years is split into 7 subsets each consisting of 
4 randomly selected years. While 6 of these subsets are used as learning 
data the 7th is the testing data set for which predictions are made and 
evaluated. This is done by calculating the CRPS value for each of the 
observations in the testing data set and then averaging them. Each of the 
7 parts is selected once as testing data which leads to 7 different settings 
of which again the average values are stored. This whole procedure is repeated 
10 times. The resulting skill scores are illustrated in 
Figure~\ref{boxplots_crps_axams} using the EMOS model as the reference model 
represented by the horizontal line at height 0. This CRPS skill score boxplot 
shows that, in this setting, both GAMs and the distributional forest perform 
better than the EMOS model. While the two GAMs lead to an improvement of 
around 4 percent with the boosted version being slightly ahead, the 
distributional forest even reaches an improvement rate close to 6 percent.
<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams_7x10.rda")){
  load("rain_Axams_7x10.rda")
} else {
  source("rain_cross_7x10.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)
